---
output: 
  html_document: 
    keep_md: yes
---

# Reproducible Research: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE }
library(ggplot2)
library(scales)
library(Hmisc)
```

## Loading and preprocessing the data
##### 1. Load the data 
```{r, results='markup', warning=TRUE, message=TRUE}
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
data <- read.csv('activity.csv')
```

-----

## What is mean total number of steps taken per day?
```{r}
stepsPerDay <- tapply(data$steps, data$date, sum, na.rm=TRUE)
```

##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsPerDay, xlab='Total steps per day', ylab='Frequency using binwith 500', binwidth=500)
```

##### 2. Calculate the mean and median 
```{r}
stepsPerDayMean <- mean(stepsPerDay)
stepsPerDayMedian <- median(stepsPerDay)
```
* Mean: `r stepsPerDayMean`
* Median:  `r stepsPerDayMedian`

-----

## What is the average daily activity pattern?
```{r}
averageStepsPerTimeBlock <- aggregate(x=list(meanSteps=data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meanSteps)) +
    geom_line() +
    xlab("5-minute interval") +
    ylab("Average number of steps taken") 
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
mostSteps <- which.max(averageStepsPerTimeBlock$meanSteps)
timeMostSteps <-  gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", averageStepsPerTimeBlock[mostSteps,'interval'])
```

* Most Steps at: `r timeMostSteps`

----

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
numMissingValues <- length(which(is.na(data$steps)))
```

* Number of missing values: `r numMissingValues`


##### 3. Create a new dataset.

```{r}
dataImputed <- data
dataImputed$steps <- impute(data$steps, fun=mean)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r}
stepsPerDayImputed <- tapply(dataImputed$steps, dataImputed$date, sum)
qplot(stepsPerDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsPerDayMeanImputed <- mean(stepsPerDayImputed)
stepsPerDayMedianImputed <- median(stepsPerDayImputed)
```
* Mean (Imputed): `r stepsPerDayMeanImputed`
* Median (Imputed):  `r stepsPerDayMedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?



```{r}
dataImputed$dateType <-  ifelse(as.POSIXlt(dataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r}
averagedDataImputed <- aggregate(steps ~ interval + dateType, data=dataImputed, mean)
ggplot(averagedDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```

